\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{xcolor}
\usepackage{enumitem}


\usepackage[colorlinks=true,                % ссылки будут цветными, а не в рамке
            linkcolor=blue,                  % цвет внутренних ссылок (на главы, формулы и т.д.)
            urlcolor=blue,                    % цвет ссылок на интернет-страницы
            citecolor=blue,                    % цвет ссылок на библиографию
            pdfborder={0 0 0}                  % убирает рамку (на всякий случай)
           ]{hyperref}

% Custom commands for consistent formatting
\newcommand{\skill}[1]{\textbf{#1}}
\newcommand{\tableref}[1]{Table~\ref{#1}}
\newcommand{\secref}[1]{Section~\ref{#1}}
\newcommand{\pageRef}[1]{[→ p. #1]}

\title{Mythras: Game System \\ \large Reference Document (Action, Time, Movement \& Core Mechanics)}
\author{}
\date{}

\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt}

\begin{document}

\maketitle

\tableofcontents
\newpage

\section{ACID}
\label{sec:acid}

Acids come from many different sources and are used extensively in alchemical research. All acids are  classified as either Weak, Strong or Concentrated. 

A splash or spray of acid lasts only for a few \textcolor{red}{Combat Rounds} before it loses potency. Immersion in a considerable volume of acid inflicts the damage every round until the victim or location is removed and treated.

For generic acids it is assumed that armour protects against the effects of acid but does not stop it, its Armour Points being reduced by the acid's damage until it reaches zero hit points, at which point the damage is then transferred to the hit location it protected. Armour reduced to zero Armour Points is considered useless as the acid burns through bindings, straps and joints.

\begin{table}[h]
\centering
\caption{Acid Table}
\label{tab:acid}
\begin{tabular}{lll}
\toprule
\textbf{Acid Type} & \textbf{Damage} & \textbf{Duration} \\
\midrule
Weak    & 1d2 & 1 Combat Round \\
Strong  & 1d4 & 1d2 Combat Rounds \\
Concentrated & 1d6 & 1d3 Combat Rounds \\
\bottomrule
\end{tabular}
\end{table}


\section{ACTION, TIME AND MOVEMENT}


\subsection{Traveling}
Typical distances that are covered during travel are listed below:

\begin{table}[h]
\centering
\caption{Time Travel Table}
\label{tab:strategic-travel}
\begin{tabular}{ll}
\toprule
\textbf{Travel Example} & \textbf{Distance Covered (km) per Day} \\
\midrule
Walking & 30 per Day \\
Horseback at casual speed & 60 per Day \\
Wagon at casual speed & 15 per Day \\
Open Sea, favourable conditions & 150-300 (24h) \\
Open Sea, unfavourable conditions & 0-60 (24h) \\
Coast or River, favourable & 30-60 per Day \\
Coast or River, unfavourable & 0-30 per Day \\
\bottomrule
\end{tabular}
\end{table}

If characters need to increase the distances given in the Time Travel Table, then they can effectively add half again to the distance travelled in the Strategic Time period but gain an enduring level of 
\hyperref[sec:fatigue]{Fatigue} as a result. Naturally an appropriate skill roll (Drive, Ride, Athletics, and so on) also needs to be made to successfully increase the distance covered.


\subsection{Movement}
\label{sec:movement}
Movement is broken down into three 'gaits' which are Walking, Running and Sprinting.
\begin{itemize}[leftmargin=*]
    \item \textbf{Walking} is the average speed a member of a particular species ambles along at when in no particular hurry. This is normally referred to as a creature's base Movement Rate.
    \item \textbf{Running} is a trot or jog, at a speed which can be maintained over long periods. Maximum running speed is triple Movement Rate, which may vary according to the bonus granted by Athletics skill \secref{sec:skills}.
    \item \textbf{Sprinting} is flat out movement at top speed, which can only be maintained for brief periods. Peak sprinting speed is five times Movement Rate, which again may be increased according to whatever bonus is granted from Athletics skill.
\end{itemize}



\begin{table}[h]
\centering
\caption{Comparative Movement Table (in meters)}
\label{tab:comparative-movement}
\begin{tabular}{lcccccc}
\toprule
\textbf{Time Period} & \textbf{Mov 4} & \textbf{Mov 6} & \textbf{Mov 8} & \textbf{Mov 10} & \textbf{Mov 12} & \textbf{Mov 14} \\
\midrule
Combat Round & 4m & 6m & 8m & 10m & 12m & 14m \\
Minute & 48m & 72m & 96m & 120m & 144m & 168m \\
Hour & 2.9km & 4.3km & 5.8km & 7.2km & 8.6km & 10.1km \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Moving in Armour}
\label{sec:armor-movement}
Worn armour acts against character Movement Rates, and certain kinds of actions such as swimming or climbing. The armour's \hyperref[sec:armour]{Initiative Penalty} is applied to Movement of different kinds in the following ways:
\begin{itemize}[leftmargin=*]
    \item \textbf{Walking:} Armour does not interfere with walking movement, although it can increase the \hyperref[sec:fatigue]{Fatigue level}.
    \item \textbf{Running or Sprinting:} Subtract the Armour Penalty from the running and sprinting speed.
    \item \textbf{Swimming:} Take the character's \textcolor{red}{swimming speed}  divide by two (rounding up), and subtract the Armour Penalty. If the result is zero, the character cannot move, and barely keeps himself afloat. If the result is negative, then the character sinks.
    \item \textbf{Climbing rough surface:} Half the Armour Penalty (rounded up) is subtracted from the base Movement Rate.
    \item \textbf{Climbing a steep surface:} The Armour Penalty is subtracted from the base Movement Rate.
    \item \textbf{Climbing a sheer surface:} Double the Armour Penalty is subtracted from the base Movement Rate.
    \item \textbf{Jumping:} Reduce the distance in metres \textcolor{red}{the character can jump}  by half the Armour Penalty (rounded up). For standing jumps this impairment is halved.
\end{itemize}


\section{AGEING}
\label{sec:ageing}
All characters age, and with age come certain consequences. The signs of ageing start at Early Middle Age (40 years for humans). As a character passes into a new Ageing Band he must make both an Endurance roll and a Willpower roll at the grades noted. If a roll is failed then he experiences Ageing Effects as shown in the Ageing Effects table.

Each characteristic affected by ageing reduces by 1d3 points. If any characteristic is reduced to zero from ageing, the character dies due to his terminal frailty.

\begin{table}[h]
\centering
\caption{Age Bands}
\label{tab:age-bands}
\begin{tabular}{ll}
\toprule
\textbf{Age Band} & \textbf{Roll Grade} \\
\midrule
Early Middle Age (40-49) & Easy \\
Middle Age (50-59) & Standard \\
Late Middle Age (60-69) & Hard \\
Old Age (70-79) & Formidable \\
Advanced Old Age (80-89) & Herculean \\
Doddering (90+) & Hopeless \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[h]
\centering
\caption{Ageing Effects Table}
\label{tab:ageing-effects}
\begin{tabular}{cll}
\toprule
\textbf{1d6} & \textbf{Physical (Failed Endurance)} & \textbf{Mental (Failed Willpower)} \\
\midrule
1-2 & STR & INT \\
3-4 & CON & POW \\
5-6 & DEX & CHA \\
\bottomrule
\end{tabular}
\end{table}

\section{ASPHYXIATION, DROWNING AND SUFFOCATION}
\label{sec:asphyxiation}
Characters can hold their breath for a number of seconds equal to their Endurance skill. However the character must be prepared (filling the lungs with as much air as possible); if not, then the period is halved if the character was in a passive situation, or reduced to one fifth if the character was engaged in strenuous activity.

Once the period of held breath is over, characters must make an Endurance roll:
\begin{itemize}[leftmargin=*]
    \item If the roll is a critical success, no further deterioration occurs.
    \item If the roll is successful, the character accrues an extra level of \hyperref[sec:fatigue]{Fatigue}.
    \item If the roll fails, the character sustains 1d2 extra levels of \hyperref[sec:fatigue]{Fatigue} that round.
    \item If the roll is fumbled, the character sustains 1d3 extra levels of \hyperref[sec:fatigue]{Fatigue} that round.
\end{itemize}
Without aid, death from asphyxiation is usually swift. If the asphyxiation ends before the character dies, they recover Fatigue levels lost to suffocation relatively quickly; regaining one level per minute.


\section{CHARACTER IMPROVEMENT}
\label{sec:improvement}
The mechanism for character improvement is the Experience Roll. Games Masters dispense Experience Rolls at an appropriate juncture in the campaign. The number of Experience Rolls given in any one sitting should be between two and four.

\subsection{Increasing Existing Skills}
\label{sec:skill-improvement}
Any skill on the character sheet, Standard or Professional, can be increased by spending one Experience Roll.
\begin{itemize}[leftmargin=*]
    \item The player rolls 1d100 and compares it to the skill being increased. The character's INT is added to the roll.
    \item If the number rolled is equal to or greater than the skill being improved it increases by 1d4 + 1\%.
    \item If the number rolled is less than the skill selected, the skill still increases but only by 1\%.
    \item If a character fumbled any skill during the course of the preceding session(s), the fumbled skill gains a free increase of 1\%.
\end{itemize}

\subsection{Increasing Characteristics}
\label{sec:characteristic-improvement}
Characteristics can, like skills, be improved through Experience Rolls, which represent training regimes. Such increases are artificial boosts which normally atrophy after the training exercises cease.

To achieve and maintain characteristic increases requires that a character reduce his regular allotment of Experience Rolls by one or more points. Each Experience Roll sacrificed in this manner boosts the trained characteristic by one tenth of its rolled \textcolor{red}{species maximum}.

No matter how much training is undertaken, no characteristic can exceed its species maximum. SIZ is the exception to the above rules. It cannot be increased through mundane means.

\subsection{Increasing Passions}
\label{sec:passions-improvement}
As described under Passions, the value of a passion may be increased with Experience Rolls in exactly the same way as a skill.

\subsection{Training}
\label{sec:training}

Skills can be improved without expending Experience Rolls, through help of a mentor; either a trainer or a teacher. A trainer must have at least 20\% more than the character in the skill being trained, and the degree of improvement rests on the difference.

Characters must spend one full week in training to benefit from a training increase. At the end of the training period the skill being trained improves by the die roll indicated on the Training Chart, with any modifiers due to the Teaching skill.

\begin{table}[h]
\centering
\caption{Training Chart}
\label{tab:training}
\begin{tabular}{ll}
\toprule
\textbf{Difference} & \textbf{Improvement} \\
\midrule
Less than 20\% & 1d2 \\
21-30\% & 1d2 \\
31-40\% & 1d3 \\
41-50\% & 1d4 \\
51-60\% & 1d6 \\
61-70\% & 1d6+1 \\
71-80\% & 1d6+2 \\
81-90\% & 1d6+3 \\
91-100\% & 1d6+4 \\
Each 10\% & +1 \\
\bottomrule
\end{tabular}
\end{table}

\subsubsection{Teachers}
\label{sec:teachers}
A teacher is a professional trainer who possesses the Teach skill, using it to improve the quality of his tuition. They utilise the same Training Chart above, but modify the amount increased according to the result of a Teach roll:
\begin{itemize}[leftmargin=*]
    \item Critical Success: Skill improvement increases by two Training steps
    \item Success: Skill improvement increases by one Training step
    \item Failure: No effect on improvement
    \item Fumble: Skill improvement decreases by one Training step, which may result in the character gaining no improvement through the training
\end{itemize}

\section{DISEASE AND POISON}
\label{sec:disease-poison}
All diseases and poisons manifest a number of traits important to their effects.

\subsection{Disease and Poison Traits}
\begin{itemize}[leftmargin=*]
    \item \textbf{Application:} The method of how the disease or poison is introduced into the victim (Ingestion, Inhaled, Contact, Injected).
    \item \textbf{Potency:} The virulence of the disease or poison. This value is set against an appropriate resisting skill (usually Endurance or Willpower) in an opposed roll.
    \item \textbf{Resistance:} How the disease or poison is resisted - either Endurance or Willpower.
    \item \textbf{Onset Time:} The delay before the disease or poison takes effect.
    \item \textbf{Duration:} How long a disease or poison's Conditions last.
    \item \textbf{Conditions:} Every toxin has one or more Conditions with specific effects.
    \item \textbf{Antidote/Cure:} If the toxin can be treated it will be noted here.
\end{itemize}

\begin{longtable}{lp{10cm}}
\caption{Disease and Poison Conditions}
\label{tab:disease-conditions} \\
\toprule
\textbf{Condition} & \textbf{Effects} \\
\midrule
\endfirsthead
\multicolumn{2}{c}{\tablename\ \thetable\ -- continued from previous page} \\
\toprule
\textbf{Condition} & \textbf{Effects} \\
\midrule
\endhead
\bottomrule
\endfoot
Agony & Victim is hindered by intense pain. Any skill involving use of the affected area must also be less or equal to the character's Willpower, otherwise the attempt fails. \\
Asphyxiation & Victim suffers \hyperref[sec:asphyxiation]{asphyxiation}  – he collapses incapacitated, unable to breathe. \\
Bleeding & Victim suffers from either internal bleeding or surface haemorrhaging which leads to blood loss. \\
Blindness & Victim becomes blind. \\
Confusion & Victim cannot use any knowledge, communication or magic skill. \\
Contagious & Victim can transfer the poison or disease by touch. \\
Deafness & Victim loses his hearing. \\
Death & Victim collapses incapacitated, and dies after a number of rounds equal to his CON characteristic. \\
Dumbness & Victim's vocal chords are paralysed, preventing verbal communication. \\
Exhaustion & Victim gains an extra level of \hyperref[sec:fatigue]{Fatigue}. \\
Fever & Victim's body temperature fluctuates wildly. All skills suffer a difficulty grade of Hard. \\
Hallucinations & Victim experiences delusions and cannot differentiate between real and imaginary experiences. \\
Maiming & Victim suffers a permanent loss of 1 Hit Point in the location(s) affected. \\
Mania & Victim is driven to follow some compulsion. \\
Nausea & Victim cannot eat, and must roll against his Endurance every time he performs a stressful physical action to avoid being physically sick. \\
Paralysis & Victim is unable physically to move. \\
Sapping & Victim has their Magic Points reduced. \\
Unconscious & Victim loses consciousness for a period specified. \\
\end{longtable}

\subsection{Sample Diseases}
\label{sec:sample-diseases}
\textbf{Pyrohoxia}
\begin{itemize}[leftmargin=*]
    \item Application: Injected
    \item Potency: 80
    \item Resistance: Endurance
    \item Onset time: 1d6+6 days
    \item Duration: 1 week + 1d3 days
    \item Conditions: Disease manifests with the start of Hallucinations. After one week they also start to suffer from Mania (Fire). If the victim somehow survives to the conclusion of the disease, they then suffer Death.
\end{itemize}

\textbf{Red Pox}
\begin{itemize}[leftmargin=*]
    \item Application: Contact
    \item Potency: 50
    \item Resistance: Endurance
    \item Onset time: 1d6+6 hours
    \item Duration: 1d6+3 days
    \item Conditions: Once the disease manifests the victim becomes Contagious. 1 day later they begin to suffer Fever and Bleeding. Each day the pustules weep blood the victim loses one level of Fatigue.
\end{itemize}

\textbf{Soul Leech}
\begin{itemize}[leftmargin=*]
    \item Application: Inhaled
    \item Potency: 65
    \item Resistance: Endurance
    \item Onset time: 1d6 days
    \item Duration: 1d3+3 weeks
    \item Conditions: Sapping. Each week the victim reduces his Magic Points attribute by 1d6 points.
\end{itemize}

\subsection{Sample Poisons}
\label{sec:sample-poisons}
\textbf{Cobra Venom}
\begin{itemize}[leftmargin=*]
    \item Application: Contact (eyes) or Injected
    \item Potency: 75
    \item Resistance: Endurance
    \item Onset time: Instant if spat in eyes, 1d6+4 minutes if bitten
    \item Duration: 1d3+3 days
    \item Conditions: If spat into the eyes both Agony and Blindness are instant. Bitten victims also begin with Agony but can struggle along until 1d6+6 hours after the bite when Asphyxiation strikes. Survivors will then suffer Necrosis, losing 1 Hit Point per day from the location bitten.
\end{itemize}

\textbf{Lotus Dust}
\begin{itemize}[leftmargin=*]
    \item Application: Inhaled
    \item Potency: 90
    \item Resistance: Willpower
    \item Onset time: 1d3 rounds
    \item Duration: Instantaneous in its natural state, 2d6 hours if smoked
    \item Conditions: Death if natural pollen. Paralysis and Hallucinations if smoked.
\end{itemize}

\textbf{Sleeping Draught}
\begin{itemize}[leftmargin=*]
    \item Application: Ingested
    \item Potency: 60
    \item Resistance: Willpower
    \item Onset time: 1d6+4 minutes
    \item Duration: 1d6+3 hours
    \item Conditions: Unconsciousness.
\end{itemize}

\section{ENCUMBRANCE}
\label{sec:encumbrance}
Encumbrance represents both the mass and bulk of an item. Items that have a zero ENC value are, on their own, inconsequential; however consider that 20 zero ENC items equal 1 ENC.

\subsection{Encumbrance Capacity}
\label{sec:enc-capacity}
Characters can carry a total ENC equal to their STR × 2 with relative ease. Everyday clothing does not contribute to this capacity, but armour does.

If the total ENC borne exceeds STR × 2 then the character is considered to be \textbf{Burdened}. This has the following effects:
\begin{itemize}[leftmargin=*]
    \item Skills using STR or DEX as part of their base become one grade harder
    \item Base Movement Rate drops by 2 metres, and the character can no longer sprint
    \item Carrying the load counts as \hyperref[tab:physical-effort]{Medium activity} for Fatigue purposes 
\end{itemize}

If the borne ENC exceeds STR × 3 then the character is considered to be \textbf{Overloaded}:
\begin{itemize}[leftmargin=*]
    \item Skills using STR or DEX as part of their base become two grades harder
    \item Base Movement Rate drops to half normal, and the character cannot move faster than a walk
    \item Carrying the load counts as \hyperref[tab:physical-effort]{Strenuous activity} for Fatigue purposes
\end{itemize}

Characters cannot carry a total ENC more than their STR × 4.

\subsection{Armour ENC}
\label{sec:armor-enc}
The amount armour counts towards Encumbrance Capacity depends on whether it is worn or carried.
\begin{itemize}[leftmargin=*]
    \item When worn, only half the total ENC of the armour is counted towards Encumbrance
    \item When carried, the full ENC value of the armour is counted towards Encumbrance
\end{itemize}


\section{FALLING}
\label{sec:falling}
The amount of damage suffered in a fall depends on the distance of the drop. Armour points do not reduce falling damage.

\begin{table}[h]
\centering
\caption{Falling Distance Table}
\label{tab:falling}
\begin{tabular}{ll}
\toprule
\textbf{Distance Fallen} & \textbf{Damage Taken} \\
\midrule
1m or less & No damage \\
2m to 5m & 1d6 to one random location \\
6m to 10m & 2d6 to two random locations \\
11m to 15m & 3d6 to three random locations \\
16m to 20m & 4d6 to four random locations \\
Each +5m & +1d6 damage \\
\bottomrule
\end{tabular}
\end{table}

Damage is rolled separately for each location; it is not spread among them.

\subsection{Size Modifiers}
\label{sec:falling-size}
Creatures of smaller size suffer less from a fall. Creatures of larger size suffer more from a fall. For every 10 points above SIZ 20 the creature adds +1d6 to the damage.

\begin{table}[h]
\centering
\caption{Size Modifiers for Falling Damage}
\label{tab:falling-size}
\begin{tabular}{ll}
\toprule
\textbf{Creature SIZ} & \textbf{Distance Fallen Adjustment} \\
\midrule
SIZ 8 to 9 & Treat distance as 1 metre less \\
SIZ 6 to 7 & Treat distance as 3 metres less \\
SIZ 4 to 5 & Treat distance as 5 metres less \\
SIZ 2 to 3 & Treat distance as 8 metres less \\
SIZ 1 or less & Treat distance as 10 metres less \\
\midrule
\textbf{Larger Creatures} & \textbf{Damage Adjustment} \\
\midrule
Each 10 points above SIZ 20 & +1d6 damage \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Mitigation}
Acrobatics can be used to mitigate falling damage - a successful roll allows the character to treat the fall as if it were two metres shorter than it actually is. Characters falling onto soft surfaces may treat the distance they fall as halved for the purposes of damage.

\section{FATIGUE}
\label{sec:fatigue}
Fatigue measures tiredness and its incremental effects. The primary way of accruing Fatigue is by engaging in some form of physical activity.

\begin{table}[h]
\centering
\caption{Physical Effort Table}
\label{tab:physical-effort}
\begin{tabular}{lll}
\toprule
\textbf{Effort} & \textbf{How Long?} & \textbf{Skill Roll} \\
\midrule
Light & CON in hours & Very Easy (Athletics/Brawn/Endurance) \\
Medium & CON in minutes & Easy (Athletics/Brawn/Endurance) \\
Strenuous & CON in seconds & Standard (Athletics/Brawn/Endurance) \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Effects of Fatigue}
 Each level of Fatigue carries penalties for skill use, movement, Initiative and Action Points.

\begin{table}[h]
\centering
\caption{Fatigue Levels}
\label{tab:fatigue-levels}
\begin{tabular}{lllll}
\toprule
\textbf{Level} & \textbf{Skill Grade} & \textbf{Move} & \textbf{Init} & \textbf{AP} \\
\midrule
Fresh & No Penalty & — & — & — \\
Winded & Hard & -1m & — & — \\
Tired & Hard & — & -2 & — \\
Wearied & Formidable & -2m & -4 & — \\
Exhausted & Formidable & Halved & -6 & -1 \\
Debilitated & Herculean & Halved & -8 & -2 \\
Incapacitated & Herculean & Immobile & — & -3 \\
Semi-Conscious & Hopeless & \multicolumn{3}{c}{No Activities Possible} \\
Comatose & \multicolumn{4}{c}{No Activities Possible} \\
Dead & \multicolumn{4}{c}{Dead} \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Recovering from Fatigue}
Characters recover from Fatigue depending on their \textcolor{red}{Healing Rate}. The amount of complete rest needed to recover from each level of accrued Fatigue is equal to the \textcolor{red}{Recovery Period} divided by the character's \textcolor{red}{Healing Rate}.

\section{FIRES}
\label{sec:fires}
Fires damage both people and objects; how much is dependent on the intensity of the source.

\begin{table}[h]
\centering
\caption{Fire Intensity Table}
\label{tab:fire}
\begin{tabular}{llll}
\toprule
\textbf{Intensity} & \textbf{Examples} & \textbf{Time to Ignite} & \textbf{Damage} \\
\midrule
1 & Candle & 1d4 & 1d2 \\
2 & Torch & 1d3 & 1d4 \\
3 & Campfire & 1d2 & 1d6 \\
4 & Room conflagration & 1d2 & 2d6 \\
5 & Volcanic lava & Instant & 3d6 \\
\bottomrule
\end{tabular}
\end{table}

\section{HEALING}
\label{sec:healing}
Natural healing from wounds and injuries is based on the character's \textcolor{red}{Healing Rate}. The Healing Rate dictates how many Hit Points are recovered in a location depending on the injury's nature:

\begin{itemize}[leftmargin=*]
    \item \textbf{Minor Wounds:} The character recovers Hit Points at a rate equal to their Healing Rate \textit{per day}.
    \item \textbf{Serious Wounds:} The character recovers Hit Points at a rate equal to their Healing Rate \textit{per week}.
    \item \textbf{Major Wounds:} The character recovers Hit Points at a rate equal to their Healing Rate \textit{per month}.
\end{itemize}



\subsection{Healing Restrictions}
\label{sec:healing-restrictions}
There are certain restrictions on natural healing:
\begin{itemize}[leftmargin=*]
    \item The healing character cannot engage in strenuous activity: otherwise the Healing Rate is reduced by 1d3. Thus, a character recovering from even a Minor Wound could find his progress halted if he decides to engage any physical tasks that might exacerbate his injuries.
    \item Natural healing will not begin to heal a Major Wound until that location has been treated with a successful Healing roll.
    \item Non-dismembering Major Wounds which are not treated with successful Healing roll within a number of days equal to one twentieth of the Healing skill become maimed, permanently reducing the Hit Points of the location.
\end{itemize}

\subsection{Magical Healing}
\label{sec:magic-healing}
Some magic can heal the wounds suffered by a victim. Yet there are specific restrictions as to what level of wound can be treated by each spell. 

No matter how petty the healing spell or miracle, its application is always enough to stabilise any type of wound, preventing bleeding and immediate death even if it doesn't actually cure the underlying injury. Note that this only applies to gross physical trauma, not to conditions brought about by suffocation, poison, and the like.

\subsection{Permanent Injuries}
\label{sec:permanent-injuries}
Some Major Wounds, and certain poisons or diseases inflict maiming injuries. The result of this damage permanently reduces the Hit Points on that location, forever weakening it. A location maimed in this way uses the diminished Hit Point value to calculate its new Serious and Major Wound thresholds.

\begin{table}[h]
\centering
\caption{Permanent Injury Table}
\label{tab:permanent-injury}
\begin{tabular}{cp{10cm}}
\toprule
\textbf{1d3} & \textbf{Maiming Result} \\
\midrule
1 & Character permanently loses one third of the Hit Points in that location. If a limb, this represents the maiming of a hand or foot. If the head, the character loses one of his sensory organs: eye, ear, nose or tongue. Anywhere else it denotes a disfiguring scar. \\
\midrule
2 & Character permanently loses two thirds of the Hit Points in that location. A limb is maimed from the elbow or knee down. The head loses two sensory organs. Torso exhibits a gruesomely horrible scar. \\
\midrule
3 & Location is reduced to a single Hit Point. Limbs are maimed from the shoulder or hip down. The head either loses three sensory organs, half the face or the entire jaw. Chest or abdomen shows such a horrific scar or deformation nobody seeing the healed wound can comprehend how the victim survived. \\
\bottomrule
\end{tabular}
\end{table}

\section{INANIMATE OBJECTS}
\label{sec:objects}
All inanimate objects possess Armour Points and Hit Points which are used to determine resistance to damage and destruction. Armour Points reduce damage before Hit Points are affected. Once an object's Hit Points have been reduced to zero, it is useless.

\begin{table}[h]
\centering
\caption{Inanimate Objects Armour and Hit Points}
\label{tab:objects}
\begin{tabular}{lll}
\toprule
\textbf{Object} & \textbf{Armour Points} & \textbf{Hit Points} \\
\midrule
Boulder & 10 & 40 \\
Castle gate & 8 & 120 \\
Castle wall (2m) & 10 & 250 \\
Chain/shackle & 8 & 8 \\
Club & 4 & 4 \\
Dagger & 6 & 4 \\
Hut wall (2m) & 3 & 15 \\
Iron door & 12 & 75 \\
Rope & 6 & 3 \\
War sword & 6 & 10 \\
Wooden chair & 2 & 6 \\
Wooden door (normal) & 4 & 25 \\
Wooden door (reinforced) & 6 & 30 \\
Wooden fence (2m) & 4 & 5 \\
\bottomrule
\end{tabular}
\end{table}

\section{LUCK POINTS}
\label{sec:luck}
Luck Points help differentiate heroes from the rank and file. Every character starts with a number of Luck Points. Luck Points can be used during play and, at the beginning of the next session, replenish to their usual value.

\subsection{Using Luck Points}
Only one Luck Point can be used in support of a particular action.
\begin{itemize}[leftmargin=*]
    \item \textbf{Cheat Fate:} Re-roll any dice roll that affects them. Force an opponent to re-roll.
    \item \textbf{Desperate Effort:} Gain an additional \textcolor{red}{Action Point} if exhausted.
    \item \textbf{Mitigate Damage:} Suffering a Major \textcolor{red}{Wound} may spend a Luck Point to downgrade the injury to a Serious Wound.
\end{itemize}


\section{PASSIONS}
\label{sec:passions}
A Passion is any deeply held commitment that has the capacity to influence events during play.

\subsection{Using Passions}
\begin{itemize}[leftmargin=*]
    \item To augment another skill, reflecting the depth of one's feeling. Adds 20\% of its value to a skill (if allowed).
    \item As an ability in its own right to drive choices. A standard roll is made against a Passion.
    \item To oppose other Passions - even those held by the same character. This is typically used where two Passions would conflict.
    \item As a general measure of depth of commitment.
    \item To resist psychological manipulation or magical domination, substituting for Willpower (if allowed).
\end{itemize}

\subsection{Deepening and Waning}
Passions can increase during a game independently of \hyperref[sec:improvement]{Experience roll} , based on the strength of whatever occurred to trigger the increase. Passions can also wane.

\begin{table}[h]
\centering
\caption{Deepening and Waning}
\label{tab:passion-change}
\begin{tabular}{ll}
\toprule
\textbf{Change} & \textbf{Amount} \\
\midrule
Weak & 1d10 \\
Moderate & 1d10+5 \\
Strong & 1d10+10 \\
\bottomrule
\end{tabular}
\end{table}

\section{SURVIVAL}
\label{sec:survival}
Characters may find themselves in hostile environments battling the elements: exposure, starvation, and thirst.

Each danger has a specific onset time:
\begin{itemize}[leftmargin=*]
    \item \textbf{Exposure:} Assuming suitable clothing it usually starts to take effect after CON hours.
    \item \textbf{Starvation:} Critical levels begin after a number of days equal to half CON.
    \item \textbf{Dehydration:} Begins after CON x4 hours.
\end{itemize}

Once conditions take hold characters start to accrue \hyperref[sec:fatigue]{Fatigue} levels. An Endurance roll is required at onset, which is then repeated every time the Exposure, Starvation or Dehydration Rate cycles.

\section{TRAPS}
\label{sec:traps}
Traps are described using the following traits.

\subsection{Trap Traits}
\begin{itemize}[leftmargin=*]
    \item \textbf{Purpose:} Alarm, Ensnaring, Maiming, Death.
    \item \textbf{Trigger:} The way the trap is set off.
    \item \textbf{Difficulty:} The challenge rating of the trap, equal to the value of the Mechanisms or Engineering skill which created it.
    \item \textbf{Resistance:} How the trap is resisted (Brawn, Evade, or parrying with a Combat Style).
    \item \textbf{Effect:} What happens when the trap is sprung.
\end{itemize}

\begin{table}[h]
\centering
\caption{Death Trap Damage Limits}
\label{tab:trap-damage}
\begin{tabular}{lll}
\toprule
\textbf{Maker's Skill} & \textbf{Damage} & \textbf{Size/Force} \\
\midrule
1-10\% & 1d2 & Small \\
11-20\% & 1d4 & Small \\
21-30\% & 1d6 & Medium \\
31-40\% & 1d8 & Medium \\
41-50\% & 1d10 & Large \\
51-60\% & 2d6 & Large \\
61-70\% & 1d8+1d6 & Huge \\
71-80\% & 2d8 & Huge \\
81-90\% & 1d10+1d8 & Enormous \\
91-100\% & 2d10 & Enormous \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Sample Traps}
\label{sec:sample-traps}
\textbf{Crushing Roof}
\begin{itemize}[leftmargin=*]
    \item Purpose: Death
    \item Difficulty: 80\%
    \item Resistance: Evade to dive clear. If trapped, combined Brawn vs 2d8.
    \item Effect: Once the ceiling lowers far enough, victims become pinned to the floor, and receive 2d8 damage per round to a random Hit Location (armour does not protect).
\end{itemize}

\textbf{Pitfall}
\begin{itemize}[leftmargin=*]
    \item Purpose: Ensnaring or Death
    \item Difficulty: 60\%
    \item Resistance: Evade to jump clear or a Hard Athletics roll to catch the edge.
    \item Effect: The drop inflicts 2d6 damage to a random Hit Location, armour does not protect.
\end{itemize}

\textbf{Spear Trap}
\begin{itemize}[leftmargin=*]
    \item Purpose: Maiming
    \item Difficulty: 75\%
    \item Resistance: Evade to dive aside or a Hard parry roll if wielding a shield.
    \item Effect: The spear inflicts 2d8 damage to a random Hit Location.
\end{itemize}

\section{VISIBILITY}
\label{sec:visibility}
The table below gives rough distances in metres for the visibility of man-sized objects, according to the ambient weather and quality of light.

The SIZ of what someone is trying to observe also affects its visibility, as will other factors such as obscuring undergrowth, background terrain, and possible camouflage. Assume targets with SIZ 10 or below halve the range at which they can be spotted, and that larger objects increase the range by one multiple for every 10 points of SIZ over 20.

\begin{table}[h]
\centering
\caption{Visibility Table (meters for man-sized objects)}
\label{tab:visibility}
\begin{tabular}{lcccc}
\toprule
\textbf{Conditions} & \textbf{Daylight} & \textbf{Twilight} & \textbf{Moonlight} & \textbf{Moonless Night} \\
\midrule
Clear & 500 & 300 & 50 & 25 \\
Overcast & 400 & 250 & 50 & 25 \\
Moderate Fog & 150 & 100 & 25 & 15 \\
Dense Fog & 50 & 30 & 15 & 10 \\
Wind-driven snow & 35 & 25 & 10 & 5 \\
Sand/dust storm & 20 & 15 & 10 & 5 \\
\bottomrule
\end{tabular}
\end{table}

\section{WEATHER}
\label{sec:weather}
Weather conditions can have a significant effect on the local environment. The main aspects of weather are precipitation, temperature, and wind. 

\subsection{Temperature}
\label{sec:temperature}
Temperature can vary dramatically due to climate, season, and elevation, depending on the campaign world and where scenarios are set. The following table provides a guideline for the effects of extreme temperature on characters.

Wearing suitable local clothing permits a character a grace period of their CON in hours before exposure sets in. After this point they need to seek shelter or start to suffer \hyperref[sec:fatigue]{Fatigue} loss. Wet characters shift the Exposure Rate one step cooler. Light or moderate gale force winds also shift Exposure Rate by one step cooler, whilst strong gales and storms shift it two steps.

\begin{table}[h]
\centering
\caption{Temperature Table}
\label{tab:temperature}
\begin{tabular}{llll}
\toprule
\textbf{Temp °C} & \textbf{Climate} & \textbf{Risk of Exposure} & \textbf{Exposure Rate} \\
\midrule
Below -20 & Glacial & Yes & Minutes \\
-19 – -10 & Freezing & Yes & 15 Minutes \\
-9 – 0 & Cold & Yes & Hourly \\
1 – 10 & Chill & Yes & Daily \\
11 – 20 & Cool & No & None \\
21 – 30 & Warm & No & None \\
31 – 40 & Hot & Yes & Hourly \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Precipitation}
\label{sec:precipitation}
Precipitation relates to the amount of moisture which falls out of the sky. Depending on the temperature it can range between rain, sleet and snow, with hail falling during storms. The base chance of it raining is equal to the relative humidity. The amount of rain per hour and duration of the fall can be calculated by looking up the relevant entry on the Precipitation table.

Normally precipitation has little effect on characters save to slow down travel if excessive rain causes flooding, or snow begins to drift, obscuring or blocking paths.

\begin{table}[h]
\centering
\caption{Precipitation Table}
\label{tab:precipitation}
\begin{tabular}{lllll}
\toprule
\textbf{Rel. Humidity} & \textbf{Typical Cloud Cover} & \textbf{Amount per Hour} & \textbf{Duration} & \textbf{Dehydration Rate} \\
\midrule
0-12\% & None & None & None & Hourly \\
13-25\% & Scant cloud & Very light (0-1mm) & 1d10 minutes & 2 Hours \\
26-37\% & Scattered cloud & Light (1-2.5mm) & 1d6 ×10 minutes & 3 Hours \\
38-50\% & Heavy cloud & Moderate (2.5-10mm) & 1d2 hours & 4 Hours \\
51-62\% & Slightly Overcast & Heavy (11-25mm) & 1d3 hours & 4 Hours \\
63-75\% & Moderately Overcast & Very Heavy (26-50mm) & 1d6 hours & 3 Hours \\
76-87\% & Completely Overcast & Monsoon (51-80mm) & 1d8 hours & 2 Hours \\
88-100\% & Storm Clouds & Deluge (81+mm) & 1d12 hours & Hourly \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Wind}
\label{sec:wind}
Wind, especially very strong winds, can have an adverse effect on activity. A wind's Strength (STR) is expressed in kilometres per hour of velocity. Its effect on physical skills – those involving STR or DEX – is detailed in the Skill Grade column; effects on Movement Rate in the Movement Rate column.

\begin{table}[h]
\centering
\caption{Wind Table}
\label{tab:wind}
\begin{tabular}{llll}
\toprule
\textbf{Wind STR} & \textbf{Name} & \textbf{Skill Grade} & \textbf{Movement Rate} \\
\midrule
0-15 & Calm Day/Light Breeze & Standard & Normal \\
16-30 & Moderate Breeze & Standard & Normal \\
31-45 & Strong Breeze & Standard & Two Thirds \\
46-60 & Light Gale & Hard & Two Thirds \\
61-75 & Moderate Gale & Hard & Half \\
76-90 & Strong Gale & Formidable & Half \\
91+ & Storm/Hurricane & Formidable & One Third \\
\bottomrule
\end{tabular}
\end{table}



\section{Currency and Economy}
\label{sec:currency}

The standard unit of currency in Mythras is the Silver Piece (SP). The exchange rate is:
\begin{itemize}
    \item 1 Copper Piece (CP) = 1/10 SP
    \item 1 Gold Piece (GP) = 10 SP
\end{itemize}

\subsection{Standard of Living}
\label{sec:living-standards}

Social class determines baseline income and living standards. \tableref{tab:social-class} shows the minimum income required to maintain a given social class.

\begin{longtable}[c]{|l|c|c|c|l|}
\hline
\multicolumn{5}{|c|}{\textbf{Table 1: Social Class}} \\
\hline
\textbf{Social Class} & \textbf{SP per Day} & \textbf{SP per Week} & \textbf{SP per Year*} & \textbf{Typical Home} \\
\hline
Exile & \multicolumn{3}{c|}{Lives a hand to mouth existence} & None/Lives Rough \\
\hline
Slave & 1 & 7 & 375 & Shared Shack/Hut \\
\hline
Freeman & 2 & 15 & 750 & Hut/Cabin/Rented Room \\
\hline
Gentry & 10 & 75 & 3,750 & House/Apartment \\
\hline
Aristocracy & 50 & 350 & 20,000 & Large Villa/Mansion \\
\hline
Ruling & 250 & 1,750 & 100,000 & Palaces/Estates \\
\hline
\multicolumn{5}{l}{*Depends on campaign setting. Assumes 360-day year, totals rounded up.} \\
\hline
\end{longtable}
Characters must spend their class-based income on general living expenses (food, drink, lodging, clothing, etc.). Spending less risks dropping a social class and reputational damage.



\section{Bartering and Haggling}
\label{sec:trade}

\subsection{Bartering}
\label{sec:bartering}

Bartering depends on offering something the other party needs or finds valuable in their normal environment.

\textbf{Process:}
\begin{enumerate}
    \item Character makes an offer of something.
    \item If offering a common commodity or essential service obviously beneficial to the other party, the barter succeeds automatically.
    \item If offering something of less obvious value/utility, make a Customs or relevant Culture roll.
    \item Make an Opposed Influence or Commerce roll. The other party resists with Willpower or Commerce (whichever is higher).
\end{enumerate}

\begin{longtable}[c]{|c|c|c|}
\hline
\multicolumn{3}{|c|}{\textbf{Table 3: Barter Modifier}} \\
\hline
\textbf{Customs/Culture} & \textbf{Influence Bonus} & \textbf{Commerce Bonus} \\
\hline
Critical & Easy & Very Easy \\
\hline
Success & Standard & Easy \\
\hline
Failure & Hard & Standard \\
\hline
Fumble & Formidable & Hard \\
\hline
\end{longtable}

If the character wins the Opposed roll, the offer is accepted. If they fumble, they have likely insulted the other party.

\subsection{Haggling}
\label{sec:haggling}

Haggling reduces a price to a reasonable amount. Use Insight rather than Customs to gain advantage, followed by a Differential Roll.

\textbf{Process:}
\begin{enumerate}
    \item Use Insight to gain advantage in negotiation.
    \item Differential Roll: Influence or Commerce vs. Willpower or Commerce.
    \item Outcome determined by difference in success levels.
\end{enumerate}

\begin{longtable}[c]{|c|c|c|}
\hline
\multicolumn{3}{|c|}{\textbf{Table 4: Haggling Outcome}} \\
\hline
\textbf{Difference in} & \textbf{Purchaser gains} & \textbf{Seller gains} \\
\textbf{Levels of Success} & \textbf{Advantage} & \textbf{Advantage} \\
\hline
3 & 25\% of Standard Price & 200\% of Standard Price \\
\hline
2 & 50\% of Standard Price & 150\% of Standard Price \\
\hline
1 & 75\% of Standard Price & 125\% of Standard Price \\
\hline
Equal & \multicolumn{2}{c|}{Standard Price} \\
\hline
\end{longtable}


\section{Armour}
\label{sec:armour}

Each armour type defined by:
\begin{itemize}
    \item \textbf{AP:} Armour points per location covered
    \item \textbf{ENC:} Typical weight and bulk 
    \item \textbf{Cost:} In Silver Pieces, per location
    \item \textbf{Milieu:} Typical era (P - Stone Age level, A - Bronze and Iron Ages, M - European Middle Ages, R - European Renaissance, E - 17th-18th Centuries, I - Industrial Revolution)
\end{itemize}

\begin{longtable}[c]{|l|l|c|c|c|c|c|}
\hline
\multicolumn{7}{|c|}{\textbf{Table 5: Armour}} \\
\hline
\textbf{Base} & \textbf{Construction} & \textbf{AP} & \textbf{ENC} & \textbf{Cost per} & \textbf{Suit ENC/Cost/} & \textbf{Milieu} \\
\textbf{Material} & & & & \textbf{Location} & \textbf{Armour Penalty} & \\
\hline
\multirow{4}{*}{Flexible} & Natural/Cured & 1 & 2 & 20 & 14/140/3 & Any \\
\cline{2-7}
 & Padded/Quilted & 1 & 1 & 80 & 7/560/2 & Any \\
\cline{2-7}
 & Laminated & 3 & 2 & 180 & 14/1260/3 & A-M \\
\cline{2-7}
 & Scaled & 4 & 3 & 320 & 21/2240/5 & A-R \\
\hline
\multirow{4}{*}{Rigid} & Half Plate & 5 & 4 & 500 & 28/3500/6 & A-R \\
\cline{2-7}
 & Mail & 6 & 5 & 900 & 30/6300/7 & A-R \\
\cline{2-7}
 & Plated Mail & 7 & 6 & 1400 & 42/9800/9 & M-E \\
\cline{2-7}
 & Articulated Plate & 8 & 7 & 2400 & 49/16800/10 & M-E \\
\hline
\end{longtable}


\subsection{Armour and SIZ}
\label{sec:armour-siz}

Flexible armour fits one point of SIZ more or less than designed. Rigid armour only fits the specific SIZ and frame it was designed for.

\subsection{Armour Penalty to Initiative}
\label{sec:armour-penalty}

Initiative Penalty = (Total ENC of all armour worn) ÷ 5, rounded up.

\textbf{Example:} Full suit of plated mail (ENC 6 per location × 7 locations = 42 total) → 42 ÷ 5 = 8.2, rounded up to 9 Initiative Penalty.

\section{Livestock}
\label{sec:livestock}

\begin{longtable}[c]{|l|c|c|c|}
\hline
\multicolumn{4}{|c|}{\textbf{Table 7: Livestock}} \\
\hline
\textbf{Item} & \textbf{Poor Quality} & \textbf{Reasonable} & \textbf{Superior} \\
\hline
Bull & 500 & 1000 & 3000 \\
\hline
Calf & 30 & 60 & 180 \\
\hline
Camel & 1100 & 2200 & 6000 \\
\hline
Cow & 100 & 200 & 600 \\
\hline
Fowl & 1 & 2 & 4 \\
\hline
Goat & 2 & 5 & 50 \\
\hline
Horse (cavalry/war-trained) & 3000 & 6000 & 10,000 \\
\hline
Horse (heavy draught) & 1400 & 2800 & 8000 \\
\hline
Horse (riding) & 1200 & 2500 & 7000 \\
\hline
Ox & 150 & 300 & 800 \\
\hline
Pig & 2 & 5 & 50 \\
\hline
Sheep & 2 & 5 & 50 \\
\hline
\end{longtable}

\section{Tools}
\label{sec:tools}
Tools include all the items any self-respecting adventurer can be
expected to own: rope, sacks and bags, candles, lanterns, and so on.
All items in this list are deemed to be of reasonable quality.


\begin{longtable}[c]{|l|c|c|}
\hline
\multicolumn{3}{|c|}{\textbf{Table 8: Tools}} \\
\hline
\textbf{Item} & \textbf{ENC} & \textbf{Cost} \\
\hline
Abacus & 1 & 8 SP \\
\hline
Astrolabe (Ancient onwards) & 1 & 200 SP \\
\hline
Backpack/Satchel (holds up to 20 ENC) & 1 & 5 SP \\
\hline
Bedroll & 1 & 1 SP \\
\hline
Bit \& Bridle & 1 & 15 SP \\
\hline
Block \& tackle & 1 & 15 SP \\
\hline
Bottle, glass/earthenware & — & 2 SP \\
\hline
Candle, 6 hour & 1 & 1 SP \\
\hline
Chain, 2m & 2 & 40 SP \\
\hline
Chest, Large (holds up to 100 ENC) & 5 & 80 SP \\
\hline
Chest, Small (holds up to 50 ENC) & 3 & 40 SP \\
\hline
Cooking Pot (for travel) & 2 & 3 SP \\
\hline
Craft tools (basic tools to match a handicraft) & 2 & 75 SP \\
\hline
Crop (riding) & — & 15 SP \\
\hline
Crowbar & 1 & 25 SP \\
\hline
Doorjamb & 2 & 5 SP \\
\hline
Driver's whip & — & 25 SP \\
\hline
Falconer's Kit & 1 & 30 SP \\
\hline
Feed/mount rations (per day) & 1 & 5 CP \\
\hline
Fine wine goblet (glass) & — & 8 SP \\
\hline
First aid kit (10 applications) & — & 25 SP \\
\hline
Fish hooks (20) & — & 1 SP \\
\hline
Fishing kit & 1 & 15 SP \\
\hline
Fishing Net & 4 & 10 SP \\
\hline
Flint \& tinder & — & 1 SP \\
\hline
Game Snares and Traps & 1 & 1 SP \\
\hline
Goad & 1 & 25 SP \\
\hline
Grappling hook & — & 5 SP \\
\hline
Hammer/Saw/Mallet/Chisel & 1 & 1 SP \\
\hline
Healer's kit (10 applications) & 1 & 150 SP \\
\hline
Hourglass (Ancient onwards) & 1 & 20 SP \\
\hline
Knife (tool, not weapon) & — & 5 SP \\
\hline
Ladder, rope 3m & 4 & 2 SP \\
\hline
Lantern, basic & 1 & 10 SP \\
\hline
Lock picks & — & 75 SP \\
\hline
Lockbox & 1-4 & 75-250 SP \\
\hline
Lodestone & — & 5 SP \\
\hline
Milling Stone & 2 & 8 SP \\
\hline
Mirror (hand glass) & 1 & 12 SP \\
\hline
Mug/Beaker/Dish/Plate (wood) & — & 5 CP each \\
\hline
Musical instrument & 2 & 70 SP \\
\hline
Nails or tacks (50) & — & 2 CP \\
\hline
Oil flask & 1 & 1 SP \\
\hline
Papyrus or paper sheet & — & 5 CP \\
\hline
Parchment (per sheet) & — & 1 SP \\
\hline
Pickaxe & 2 & 15 SP \\
\hline
Pitchfork & 2 & 5 SP \\
\hline
Pulley & 2 & 10 SP \\
\hline
Quill pen & — & 1 CP \\
\hline
Rope, 10m & 2 & 5 SP \\
\hline
Sack, large (holds 40 ENC) & 1 & 2 SP \\
\hline
Sack, small (holds 20 ENC) & — & 5 CP \\
\hline
Saddle and tack (riding) & 2 & 50 SP \\
\hline
Saddle and tack (war) & 3 & 150 SP \\
\hline
Saddle and tack (draught) & 4 & 40 SP \\
\hline
Scabbard/Baldric/Quiver & 1 & 5 SP \\
\hline
Scythe & 3 & 10 SP \\
\hline
Sheath (dagger) & — & 1 SP \\
\hline
Shovel & 2 & 5 SP \\
\hline
Sledgehammer & 3 & 15 SP \\
\hline
Soap (bar) & — & 5 CP \\
\hline
Spade & 2 & 5 SP \\
\hline
Tent, large (4 person) & 8 & 50 SP \\
\hline
Tent, small (1 person) & 3 & 15 SP \\
\hline
Torch & 1 & 1 CP \\
\hline
Waterskin & 1 & 5 CP \\
\hline
Whetstone & — & 1 SP \\
\hline
Writing ink (vial) & — & 5 SP \\
\hline
\end{longtable}

\section{Melee Weapons}
\label{sec:melee-weapons}

Melee weapons are governed by a series of characteristics:

\begin{itemize}
    \item \textbf{Damage:} Amount of damage delivered on successful attack
    \item \textbf{Size:} Ability to inflict and \textcolor{red}{parry} damage (S, M, L, H, E)
    \item \textbf{\textcolor{red}{Reach}:} Length and ability to hold a foe at bay (T, S, M, L, VL)
    \item \textbf{Combat Effects:} Access to \textcolor{red}{special effects} 
    \item \textbf{ENC:} Combined encumbrance and weight
    \item \textbf{AP/HP:} Armour Points and Hit Points
    \item \textbf{Traits:} Special characteristics (see weapon-traits below)
    \item \textbf{Cost:} In Silver Pieces
    \item \textbf{Milieu:} Typical era (P - Stone Age level, A - Bronze and Iron Ages, M - European Middle Ages, R - European Renaissance, E - 17th-18th Centuries, I - Industrial Revolution)
\end{itemize}

\subsection{Weapon Traits}
\label{sec:weapon-traits}

\begin{itemize}
    \item \textbf{Barbed:} If \textcolor{red}{impaled}, will cause full damage when yanked free rather than half
    \item \textbf{Defensive:} If used to \textcolor{red}{parry} an attacking weapon, that weapon is considered one size smaller
    \item \textbf{Double-Ended:} Length remaining in hand can be reversed and used as a shorter version
    \item \textbf{Entrapping:} Allows use of the \textcolor{red}{Pin} Weapon special effect without needing to roll a Critical
    \item \textbf{Flexible:} Shield is considered one size smaller when parrying
    \item \textbf{Mounted:} Must be used from the back of a moving mount
    \item \textbf{Offensive:} Cannot be used to \textcolor{red}{parry}
    \item \textbf{Passive Block:} Can \textcolor{red}{passively block} multiple Hit Locations, must be continuous to another
    \item \textbf{Ranged:} Can parry ranged weapons
    \item \textbf{Set:} Can be used to receive a foot or mounted charge
    \item \textbf{Stealth:} Must be used from ambush
    \item \textbf{Thrown:} Can be used as a ranged weapon as well as hand-to-hand
\end{itemize}

\begin{longtable}[c]{|l|c|c|c|c|c|c|c|c|}
\hline
\multicolumn{9}{|c|}{\textbf{Table 9: One-Handed Melee Weapons}} \\
\hline
\textbf{Weapon} & \textbf{Damage} & \textbf{Size} & \textbf{Reach} & \textbf{Combat Effects} & \textbf{ENC} & \textbf{AP/HP} & \textbf{Traits} & \textbf{Milieu} & \textbf{Cost} \\
\hline
Axe, battle & 1d8+1 & M & M & Bleed, Sunder & 2 & 6/10 & — & All & 75 SP \\
\hline
Axe, great & 2d6+2 & L & L & Bleed, Sunder & 3 & 6/12 & — & A-M & 175 SP \\
\hline
Axe, hand & 1d6 & M & S & Bleed & 1 & 4/8 & Thrown & All & 15 SP \\
\hline
Club & 1d6 & M & S & Stun Location & 1 & 4/8 & — & All & 5 SP \\
\hline
Dagger & 1d4+1 & S & S & Impale & — & 6/8 & — & All & 30 SP \\
\hline
Falchion & 1d8 & M & M & Bleed, Sunder & 2 & 6/10 & — & M & 100 SP \\
\hline
Flail & 1d8 & M & L & Stun Location, Sunder & 2 & 4/8 & Flexible & M & 50 SP \\
\hline
Gladius & 1d8 & M & S & Impale & 1 & 6/10 & — & A & 100 SP \\
\hline
Hatchet & 1d6 & M & S & Bleed & 1 & 4/8 & Thrown & All & 25 SP \\
\hline
Javelin & 1d8 & M & L & Impale & 2 & 4/8 & Thrown & A-M & 20 SP \\
\hline
Knife & 1d4 & S & S & Impale & — & 4/6 & — & All & 10 SP \\
\hline
Kopis & 1d8 & M & M & Bleed, Sunder & 2 & 6/10 & — & A & 125 SP \\
\hline
Mace & 1d8 & M & S & Stun Location & 2 & 6/10 & — & All & 50 SP \\
\hline
Maul & 1d10 & L & L & Stun Location & 3 & 4/12 & — & A-M & 80 SP \\
\hline
Pick & 1d6+1 & M & M & Impale & 2 & 5/10 & — & A-M & 40 SP \\
\hline
Rapier & 1d8 & M & L & Impale & 1 & 5/10 & — & R-E & 150 SP \\
\hline
Sabre & 1d8 & M & M & Bleed & 2 & 6/10 & — & M-E & 125 SP \\
\hline
Scimitar & 1d8 & M & M & Bleed & 2 & 6/10 & — & M & 100 SP \\
\hline
Shortsword & 1d6+1 & M & S & Impale & 1 & 6/10 & — & All & 60 SP \\
\hline
Sickle & 1d6 & M & S & Bleed, Impale & 1 & 4/8 & — & A-M & 15 SP \\
\hline
Spear & 1d8+1 & L & L & Impale & 2 & 4/10 & Set & All & 20 SP \\
\hline
Staff, quarter & 1d8 & M & L & Stun Location & 2 & 4/10 & Double-Ended & All & 5 SP \\
\hline
Trident & 1d8+1 & L & L & Impale & 2 & 5/10 & Barbed & A-M & 80 SP \\
\hline
Warhammer & 1d8+1 & M & M & Stun Location & 2 & 6/12 & — & A-M & 100 SP \\
\hline
\end{longtable}

\begin{longtable}[c]{|l|c|c|c|c|c|c|c|c|}
\hline
\multicolumn{9}{|c|}{\textbf{Table 10: Two-Handed Melee Weapons}} \\
\hline
\textbf{Weapon} & \textbf{Damage} & \textbf{Size} & \textbf{Reach} & \textbf{Combat Effects} & \textbf{ENC} & \textbf{AP/HP} & \textbf{Traits} & \textbf{Milieu} & \textbf{Cost} \\
\hline
Bardiche & 2d8 & H & VL & Bleed, Sunder & 4 & 6/14 & — & M & 150 SP \\
\hline
Bill & 2d6+2 & H & VL & Impale, Sunder & 4 & 5/14 & Set & M & 80 SP \\
\hline
Falx & 2d6 & L & L & Bleed & 3 & 6/12 & — & A-M & 125 SP \\
\hline
Glaive & 2d6+1 & H & VL & Bleed, Sunder & 4 & 5/14 & Set & M & 100 SP \\
\hline
Greatsword & 2d8 & H & L & Bleed, Sunder & 4 & 6/16 & — & M-E & 350 SP \\
\hline
Halberd & 2d6+2 & H & VL & Bleed, Impale, Sunder & 4 & 5/16 & Set & M & 120 SP \\
\hline
Longaxe & 2d6+2 & H & VL & Bleed, Sunder & 4 & 6/14 & — & M & 200 SP \\
\hline
Longspear & 1d10+1 & H & VL & Impale & 3 & 4/12 & Set & A-M & 40 SP \\
\hline
Longsword & 1d10 & L & L & Bleed, Sunder & 3 & 6/14 & — & M-E & 250 SP \\
\hline
Military pick & 1d10 & L & L & Impale & 3 & 5/12 & — & A-M & 70 SP \\
\hline
Pike & 1d12+1 & E & VL & Impale & 5 & 4/16 & Set, Mounted & M-E & 60 SP \\
\hline
Scythe & 1d10+1 & H & VL & Bleed, Impale & 4 & 4/12 & — & A-M & 30 SP \\
\hline
Voulge & 2d6+1 & H & L & Bleed, Sunder & 4 & 5/14 & — & M & 90 SP \\
\hline
Zweihänder & 2d8+1 & E & VL & Bleed, Sunder & 5 & 6/18 & — & R-E & 450 SP \\
\hline
\end{longtable}

\section{Shields}
\label{sec:shields}

Shields are defensive weapons that can also be used offensively. Each shield is defined by:
\begin{itemize}
    \item \textbf{AP/HP:} Armour Points and Hit Points
    \item \textbf{Size:} Ability to deflect attacks (S, M, L, H)
    \item \textbf{ENC:} Encumbrance value
    \item \textbf{Combat Effects:} \textcolor{red}{Special effects} available
    \item \textbf{Traits:} Special characteristics
    \item \textbf{Cost:} In Silver Pieces
    \item \textbf{Milieu:} Typical era (P - Stone Age level, A - Bronze and Iron Ages, M - European Middle Ages, R - European Renaissance, E - 17th-18th Centuries, I - Industrial Revolution)
\end{itemize}

\begin{longtable}[c]{|l|c|c|c|c|c|c|}
\hline
\multicolumn{7}{|c|}{\textbf{Table 11: Shields}} \\
\hline
\textbf{Shield Type} & \textbf{AP/HP} & \textbf{Size} & \textbf{ENC} & \textbf{Combat Effects} & \textbf{Traits} & \textbf{Milieu} & \textbf{Cost} \\
\hline
Buckler & 6/8 & S & 1 & Stun Location & Defensive, Ranged & All & 25 SP \\
\hline
Heater & 8/12 & M & 2 & Stun Location & Defensive & M-E & 50 SP \\
\hline
Hoplon/Aspis & 8/16 & L & 3 & Stun Location & Defensive & A & 75 SP \\
\hline
Kite & 8/14 & L & 3 & Stun Location & Defensive & M & 70 SP \\
\hline
Round & 8/12 & M & 2 & Stun Location & Defensive & P-M & 40 SP \\
\hline
Scutum & 8/18 & H & 4 & Stun Location & Defensive, Passive Block & A & 100 SP \\
\hline
Target & 8/14 & M & 2 & Stun Location & Defensive & M-E & 60 SP \\
\hline
Wall shield & 8/20 & H & 5 & Stun Location & Defensive, Passive Block & A-M & 150 SP \\
\hline
\end{longtable}

\section{Ranged Weapons}
\label{sec:ranged-weapons}


Ranged weapons have very similar characteristics to Melee Weapons, although there are some differences as noted below:

\begin{itemize}
    \item \textbf{Damage:} Damage dealt on successful attack.
    \item \textbf{Damage Modifier:} Yes/No - whether user's damage modifier applies.
    \item \textbf{Force:} Impact rating (S/M/L/H/E) determining ability to overcome shields.
    \item \textbf{Range:} Three range bands in metres: Close/Effective/Long.
    \begin{itemize}
        \item \textbf{Close Range:} Choose Location special effect allowed against stationary or unaware targets without requiring a Critical.
        \item \textbf{Effective Range:} No significant modifiers on accuracy or power.
        \item \textbf{Long Range:} Damage halved, Force reduced by one step.
    \end{itemize}
    \item \textbf{Load:} Number of turns required to load or reload ammunition.
    \item \textbf{Combat Effects:} Same \textcolor{red}{special effects} as close combat weapons.
    \item \textbf{Impaling Size:} Effective size (S/M/L/H/E) of weapon or ammunition for impaling attacks.
    \item \textbf{ENC:} Combined encumbrance and weight.
    \item \textbf{AP/HP:} Armour Points and Hit Points - damage resistance and durability.
    \item \textbf{Cost:} Price in Silver Pieces.
    \item \textbf{Milieu:} Typical era (P - Stone Age level, A - Bronze and Iron Ages, M - European Middle Ages, R - European Renaissance, E - 17th-18th Centuries, I - Industrial Revolution)
\end{itemize}

Thrown javelins may use the Pin Weapon special effect against shields, without needing to roll a Critical.

\begin{longtable}[c]{|l|c|c|c|c|c|c|c|c|c|c|}
\hline
\multicolumn{11}{|c|}{\textbf{Table 12: Ranged Weapons}} \\
\hline
\textbf{Ranged Weapon} & \textbf{Damage} & \textbf{Dam. Mod} & \textbf{Force} & \textbf{Range} & \textbf{Load} & \textbf{Combat Effects} & \textbf{Impale Size} & \textbf{ENC} & \textbf{AP/HP} & \textbf{Milieu} & \textbf{Cost} \\
\hline
Atlatl & — & +1 Step & — & 10/25/75 & 1 & — & — & 1/4 & P & 10 SP \\
\hline
Blowgun & 1d4 & N & — & 10/20/30 & 2 & — & — & 1/4 & P & 30 SP \\
\hline
Bolas & 1d4 & N & — & 10/25/50 & — & Entangle & — & 2/2 & P-A & 10 SP \\
\hline
Dagger & 1d4 & Y & S & 5/10/20 & — & Impale & S & 6/8 & All & 30 SP \\
\hline
Dart & 1d4 & Y & S & 5/10/20 & — & Impale & S & 2/1 & P-A & 10 SP \\
\hline
Discus & 1d4+1 & Y & L & 5/20/40 & — & Stun Location & — & 2/3 & A & 30 SP \\
\hline
Hatchet & 1d6 & Y & S & 10/20/30 & — & Bleed & — & 14/6 & All & 25 SP \\
\hline
Heavy crossbow & 1d10 & N & H & 20/150/300 & 4 & Impale, Sunder & S & 24/8 & M-E & 350 SP \\
\hline
Javelin & 1d8+1 & Y & H & 10/20/50 & — & Impale, Pin Weapon (Shield) & M & 13/8 & A-M & 20 SP \\
\hline
Light crossbow & 1d8 & N & L & 20/100/200 & 3 & Impale & S & 14/5 & M-E & 150 SP \\
\hline
Long bow & 1d8 & Y & H & 15/125/250 & 2 & Impale & S & 14/7 & M & 200 SP \\
\hline
Net & — & N & — & 3/5/10 & — & Entangle & — & 32/20 & All & 20 SP \\
\hline
Recurve bow & 1d8 & Y & H & 15/125/250 & 2 & Impale & S & 14/8 & A-M & 225 SP \\
\hline
Short bow & 1d6 & Y & L & 15/100/200 & 2 & Impale & S & 14/4 & P-M & 75 SP \\
\hline
Shortspar & 1d8 & Y & L & 10/15/30 & — & Impale & M & 14/5 & All & 20 SP \\
\hline
Sling & 1d8 & N & L & 10/150/300 & 3 & Stun Location & — & 1/2 & P-M & 5 SP \\
\hline
Staff sling & 2d6 & N & E & 5/25/50 & 4 & Stun Location & — & 23/6 & A-M & 20 SP \\
\hline
Stone/Rock & 1d3 & Y & S & 5/10/20 & — & Stun Location & — & — & All & — \\
\hline
Thrown dart & 1d6 & Y & M & 10/20/30 & — & Impale & S & 1/2 & P-A & 5 SP \\
\hline
Trident & 1d8 & Y & L & 10/15/30 & — & Barbed, Impale & M & 24/10 & A-M & 155 SP \\
\hline
\end{longtable}

\subsection{Ranged Weapon Ammunition}
\label{sec:ammunition}

\begin{longtable}[c]{|l|c|}
\hline
\multicolumn{2}{|c|}{\textbf{Table 13: Ammunition Costs}} \\
\hline
\textbf{Item} & \textbf{Cost} \\
\hline
Arrows (per sheaf of a dozen) & 4 SP \\
\hline
Quarrels (per sheaf of a dozen) & 3 SP \\
\hline
Lead sling bullets (per bag of twenty) & 1 SP \\
\hline
Blowgun darts (per half dozen) & 2 SP \\
\hline
Thrown darts (per sheaf of a dozen) & 2 SP \\
\hline
\end{longtable}

\section{Equipment Manufacturing and Quality}
\label{sec:manufacturing}

\subsection{Resources}
\label{sec:resources}

Access to suitable tools, materials, and workplace determines:
\begin{itemize}
    \item Whether the task is possible
    \item How long it takes to complete
    \item Quality of the output
\end{itemize}

\begin{longtable}[c]{|l|l|}
\hline
\multicolumn{2}{|c|}{\textbf{Table 14: Resource Modifiers}} \\
\hline
\textbf{Circumstances} & \textbf{Skill Grade} \\
\hline
No tools & Hopeless \\
\hline
Improvised tools & Herculean \\
\hline
Poor Quality/Incomplete tools, or lacking crucial resources & Formidable \\
\hline
\end{longtable}

\subsection{Task Rounds}
\label{sec:task-rounds}

Building or repairing something is measured by Task Rounds. This is a variable amount of time dependent on what is being done.

\begin{longtable}[c]{|l|l|l|}
\hline
\multicolumn{3}{|c|}{\textbf{Table 15: Sample Task Rounds}} \\
\hline
\textbf{Task} & \textbf{Appropriate Skill} & \textbf{Task Round Unit} \\
\hline
Repairing clothing & Craft (Needlework) & 5 minutes – 1 hour \\
\hline
Forging metal & Craft (Blacksmith) & 5 minutes – 30 minutes \\
\hline
Forging a sword & Craft (Weaponsmith) & 1 day – 5 days \\
\hline
Erecting a wooden or stone structure & Engineering & 6 months – several years \\
\hline
Sharpening a blade & Mechanisms or Craft (Weaponsmith) & 3 to 5 minutes \\
\hline
\end{longtable}

Completing the task requires a roll against the relevant skill. Note the level of success:

\begin{itemize}
    \item Critical: +50\%
    \item Success: +25\%
    \item Failure: +0\%
    \item Fumble: -25\%
\end{itemize}

Continue making skill rolls, each taking one Task Round, until 100\% has been achieved - or until time, resources, and budget are exhausted. Reaching 100\% means the final result is of an acceptable standard. Anything less is below par, and may be flawed in some way.

With unlimited time it is possible to exceed 100\%. The percentage achieved at the end of the time available indicates the overall quality of the product or task. Exceptional quality offers enhancements to the product.

\subsection{Enhancements}
\label{sec:enhancements}

If a project reaches 100\% in four or fewer Task Rounds, the maker has the option to continue refining it. To do this he continues to work on it for a further Task Round and, depending on the result of the skill roll for the Task Round, may enhance it:

\begin{itemize}
    \item Critical Success: +2 Enhancements
    \item Success: +1 Enhancement
    \item Failure: No Enhancements
    \item Fumble: Reduce Quality by 25\%
\end{itemize}


\begin{longtable}[c]{|c|l|c|}
\hline
\multicolumn{3}{|c|}{\textbf{Table 16: Item Quality}} \\
\hline
\textbf{Total Score} & \textbf{Result} & \textbf{Enhancements} \\
\hline
150+ & Excellent & 2 \\
\hline
125-149 & Good & 1 \\
\hline
100-124 & Acceptable & 0 \\
\hline
75-99 & Flawed & \multicolumn{1}{p{8cm}|}{The product will break or wear out 25\% faster than normal; or its Hit Points and value are reduced by one quarter reflecting its flawed durability.} \\
\hline
50-74 & Poor & \multicolumn{1}{p{8cm}|}{The product will break or wear out 50\% faster than normal; or its Hit Points and value are reduced by half reflecting its flawed durability.} \\
\hline
25-49 & Shoddy & \multicolumn{1}{p{8cm}|}{The product will break or wear out 75\% faster than normal; or its Hit Points and value are reduced by three quarters reflecting its shoddy durability.} \\
\hline
0-24 & Unusable & \multicolumn{1}{p{8cm}|}{The product is simply of no use whatsoever.} \\
\hline
\end{longtable}


\begin{longtable}[c]{|l|p{10cm}|}
\hline
\multicolumn{2}{|c|}{\textbf{Table 17: Enhancements}} \\
\hline
\textbf{Enhancement} & \textbf{Effect} \\
\hline
Desirable & The item's value is doubled. \\
\hline
Durable & The item's Hit Points are increased by one. \\
\hline
Resilient & The item's Armour Points are increased by one. \\
\hline
Efficient & The item is of such quality that it confers a 5\% bonus to the skill required to use it. \\
\hline
Effective & If a weapon, it gains +1 to its Damage rating. Other items reduce ENC value by 1d2 points. \\
\hline
\end{longtable}

\section{Siege Weapons}
\label{sec:siege-weapons}

The following weapons do not use handedness, damage bonuses or ENC being beyond such issues. The Size of their projectiles is rated as Beyond Enormous (BE), meaning that no shield will protect against the force of such impacts. Due to the nature of the weapons, Range only has two values; Effective Range and Long Range.


\begin{itemize}
    \item \textbf{Reloading time} is measured in Combat Rounds, not Turns.
    \item \textbf{Crew:} This is the minimum and maximum crew required to operate the weapon. Each extra person above the minimum value reduces the Load time by 1.
    \item \textbf{Siege Weapon Combat Effects} are included for attacks against large or colossal creatures. They are superfluous against fortifications.
\end{itemize}

\begin{longtable}[c]{|l|c|c|c|c|c|c|c|}
\hline
\multicolumn{8}{|c|}{\textbf{Table 14: Siege Weapons}} \\
\hline
\textbf{Weapon} & \textbf{Damage} & \textbf{Force} & \textbf{Range (m)} & \textbf{Load} & \textbf{Crew} & \textbf{Combat Effects} & \textbf{Cost} \\
\hline
Ballista & 3d10 & BE & 250/500 & 3 rounds & 2-4 & Impale & 1000 SP \\
\hline
Catapult, light & 4d10 & BE & 150/300 & 5 rounds & 3-6 & Sunder & 2000 SP \\
\hline
Catapult, heavy & 6d10 & BE & 200/400 & 8 rounds & 4-8 & Sunder & 4000 SP \\
\hline
Trebuchet & 8d10 & BE & 300/600 & 12 rounds & 6-12 & Sunder & 8000 SP \\
\hline
Ram, battering & 2d10 & BE & — & — & 4-10 & Sunder & 500 SP \\
\hline
Siege tower & — & — & — & — & 10-30 & — & 5000 SP \\
\hline
\end{longtable}

\section{Vehicles}
\label{sec:vehicles}

Most vehicles are drawn by one or more animals, and move at the animal's rate. The number of animals required, for the animals to maintain their standard movement rate, is noted in the Draw column of the vehicles table. Reduce the movement proportionately if fewer animals are used. Generally drivers will not exceed the required number of animals because this decreases control, stability, and integrity.

The amount a vehicle can carry in terms of driver(s), passengers, and cargo (total ENC) is indicated in the Load column. Again, drivers will refuse to exceed these load amounts because control and stability becomes more difficult.

\begin{longtable}[c]{|l|c|c|c|c|c|}
\hline
\multicolumn{6}{|c|}{\textbf{Table 15: Vehicles}} \\
\hline
\textbf{Vehicle} & \textbf{Draw} & \textbf{Load (ENC)} & \textbf{Movement} & \textbf{AP/HP} & \textbf{Milieu} & \textbf{Cost} \\
\hline
Chariot, light & 2 horses & 100 & As horses & 4/20 & A-M & 300 SP \\
\hline
Chariot, heavy & 4 horses & 200 & As horses & 6/30 & A-M & 600 SP \\
\hline
Cart & 1-2 oxen/horses & 300 & Half draught animal & 4/25 & All & 100 SP \\
\hline
Wagon & 2-4 oxen/horses & 800 & Half draught animal & 6/40 & All & 300 SP \\
\hline
Wagon, covered & 2-4 oxen/horses & 600 & Half draught animal & 6/35 & All & 400 SP \\
\hline
Sled & 2-4 oxen/horses & 400 & As draught animal & 3/20 & P-M & 50 SP \\
\hline
Canoe & 1-2 paddlers & 100 & Paddler's swim rate & 2/15 & All & 30 SP \\
\hline
Rowboat & 2-4 rowers & 400 & Rower's swim rate & 3/25 & All & 100 SP \\
\hline
Sailing ship, small & Wind & 10,000 & 5-10 km/h & 8/200 & A-M & 5000 SP \\
\hline
Sailing ship, large & Wind & 50,000 & 5-10 km/h & 10/400 & A-M & 20,000 SP \\
\hline
Galley & 20-100 rowers & 30,000 & 3-7 km/h & 8/300 & A-M & 15,000 SP \\
\hline
\end{longtable}
\end{document}
